---
title: "ids"
author: "Rich FitzJohn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ids}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

``` {r echo=FALSE, results="hide"}
knitr::opts_chunk$set(error=FALSE)
human_no <- function(x) {
  s <- log10(floor(x + 1))
  p <- c(0, thousand=3, million=6, billion=9, trillion=12)
  i <- s > p
  j <- max(which(i))
  str <- names(p)[j]
  if (nzchar(str)) {
    paste(signif(x / 10^p[[j]], 3), str)
  } else {
    as.character(x)
  }
}
```

The `ids` package provides randomly generated ids in a number of
different forms with different readability and sizes.

## Random bytes

The `random_id` function generates random identifiers by generating
`bytes` random bytes and converting to hex.  Rather than use R's
random number stream we use `openssl` here.
``` {r }
ids::random_id()
```

All `ids` functions take `n` as the first argument to be the number
of identifiers generated:
``` {r }
ids::random_id(5)
```

The default here is 16 bytes (so 2^128 = 3.4e38 combinations).  You
can make these larger or smaller:
``` {r }
ids::random_id(5, 8)
```

If `NULL` is provided as `n`, then a generating function is returned:
``` {r }
f <- ids::random_id(NULL, 8)
f
```

This function sets all arguments except for `n`
``` {r }
f()
f(4)
```

## UUIDs

The above look a lot like UUIDs but they are not actually UUIDs.
The `uuid` package provides real UUIDs generated with libuuid, and
the `ids::uuid` function provides an interface to that:
``` {r }
ids::uuid()
```

As above, generate more than one UUID:
``` {r }
ids::uuid(4)
```

Generate time-based UUIDs:
``` {r }
ids::uuid(4, use_time=TRUE)
```

and optionally drop the hyphens:
``` {r }
ids::uuid(5, drop_hyphens=TRUE)
```

## Adjective animal

Generate (somewhat) human readable identifiers by combining one or
more adjectives with an animal name.
``` {r }
ids::adjective_animal()
```

The list of adjectives and animals comes from
[gfycat.com](http://gfycat.com), via
https://github.com/a-type/adjective-adjective-animal

Generate more than one identifier:
``` {r }
ids::adjective_animal(4)
```

Use more than one adjective for very long idenfiers
``` {r }
ids::adjective_animal(4, 3)
```

In addition to snake_case, the default, the punctuation between
words can be changed to:

kebab-case:
``` {r }
ids::adjective_animal(1, 2, style="kebab")
```

dot.case:
``` {r }
ids::adjective_animal(1, 2, style="dot")
```

camel-case:
``` {r }
ids::adjective_animal(1, 2, style="camel")
```

PascalCase:
``` {r }
ids::adjective_animal(1, 2, style="pascal")
```

CONSTANT_CASE
``` {r }
ids::adjective_animal(1, 2, style="constant")
```

or with spaces, lower case:
``` {r }
ids::adjective_animal(1, 2, style="lower")
```

UPPER CASE
``` {r }
ids::adjective_animal(1, 2, style="upper")
```

Sentence case
``` {r }
ids::adjective_animal(1, 2, style="sentence")
```

Title Case
``` {r }
ids::adjective_animal(1, 2, style="title")
```

``` {r echo=FALSE, results="hide"}
n1 <- length(ids:::gfycat_animals)
n2 <- length(ids:::gfycat_adjectives)
```

There are `r n1` animal names and `r n2` adjectives so each one you
add increases the idenfier space by a factor of `r n2`.  So for 1,
2, and 3 adjectives there are `r human_no(n1 * n2)`,
`r human_no(n1 * n2^2)` and `r human_no(n1 * n2^3)` possible combinations.

This is a much smaller space than the identifiers above, but these
are more readable and memorable.

Note that here, the random nunbers are coming from R's random
number stream so are affected by \code{set.seed}.

## Random sentences

The `sentence` function creates a sentence style identifier.  This
uses the approach described by Asana on [their
blog](https://blog.asana.com/2011/09/6-sad-squid-snuggle-softly).
This approach encodes 32 bits of information (so 2^32 ~= 4 billion
possibilities) and in theory can be remapped to an integer if you
really wanted to.
``` {r }
ids::sentence()
```

As with `adjective_animal`, the case can be changed:
``` {r }
ids::sentence(2, "dot")
```

If you would rather past tense for the verbs, then pass `past=TRUE`:
``` {r }
ids::sentence(4, "snake", TRUE)
```
